Syntax-Based Concept Alignment for Domain-Specific Machine Translation

### 2021-05-09 paper/paper.tex
description of MT experiments

%TODO: position properly
\begin{table*}[h]
  \centering
  \small
  \begin{tabular}{l|lll}
  %\hline
   & \textbf{CE} & \textbf{\texttt{fast\_align} 100} & \textbf{\texttt{fast\_align} 1000}\\ \hline
   \textbf{tot. alignments} & 716 & 1440 & 1435 \\ 
   \textbf{\texttt{+} and \texttt{=}} & 536 (75\%) & 410 (28\%) & 656 (46\%)\\ 
   \textbf{\texttt{+}} & 491 (69\%) & 371 (26\%) & 590 (41\%)\\ 
  \end{tabular}
  \caption[Comparison between our grammar-based CE module and \texttt{fast\_align}]{Comparison between our grammar-based CE module and \texttt{fast\_align} on PUD data. The middle column shows the results obtained by training \texttt{fast\_align} on the 100-sentence subset fed to the extraction module. On the right, the results obtained by training \texttt{fast\_align} on the full PUD treebanks dataset, discarding the alignments obtained for sentences 101-1000 for the final comparison. Row 1 reports the total number of distinct alignments obtain. Rows 2 and 3 give the percentages of correct alignments, with and without considering sentence pair-specific correspondences as correct.}
  \label{pud_fast}
\end{table*}

@@

% PUD: incl. reason-wise?)

\subsection{MT experiments}
% metrics & methods
\subsubsection{Experimental results}

>>>>

% TODO: position properly
% TODO: add english-swedish!
\begin{table*}[h]
  \centering
  \small
  \begin{tabular}{l|lll}
  %\hline
   & \textbf{CE} & \textbf{\texttt{fast\_align} 100} & \textbf{\texttt{fast\_align} 1000}\\ \hline
   \textbf{tot. alignments} & 716 & 1440 & 1435 \\ 
   \textbf{\texttt{+} and \texttt{=}} & 536 (75\%) & 410 (28\%) & 656 (46\%)\\ 
   \textbf{\texttt{+}} & 491 (69\%) & 371 (26\%) & 590 (41\%)\\ 
  \end{tabular}
  \caption[Comparison between our grammar-based CE module and \texttt{fast\_align}]{Comparison between our grammar-based CE module and \texttt{fast\_align} on PUD data. The middle column shows the results obtained by training \texttt{fast\_align} on the 100-sentence subset fed to the extraction module. On the right, the results obtained by training \texttt{fast\_align} on the full PUD treebanks dataset, discarding the alignments obtained for sentences 101-1000 for the final comparison. Row 1 reports the total number of distinct alignments obtain. Rows 2 and 3 give the percentages of correct alignments, with and without considering sentence pair-specific correspondences as correct.}
  \label{pud_fast}
\end{table*}

@@

% TODO: PUD: incl. reason-wise?)

\subsection{MT experiments}
The second set of experiments has the objective of assessing the quality of the final output of the system we propose: GF translation lexica. 
Because we are focussing on using CA in the context of domain-specific MT, in this case we do not make use of the PUD treebanks, where sentences come from a variety of different sources, but rather of the two course plans corpora. 
For a this evaluation, we do not construct a grammar specific to the course plans domain: it is sufficient to extend the extraction grammar itself, designed for the grammar generation step to produce grammar rules corresponding to simple concepts such as basic categories, noun phrases, verb phrases etc., with preexisting syntax rules defined in the RGL, so to allow for more variation. % TODO: is RGL mentioned above?

The basic idea behind the experiments whose results are presented in the following section is to automatically translate a set of English sentences to Italian and Swedish, ask native speakers of the target languages to produce a set of reference translations, and compare them to the original machine-generated ones by computing the BLEU score.

Due to the small size of the datasets and, as a consequence, the low coverage of the extracted lexicon, we generate the sentences to translate directly in the GF shell, making use of its random AST generation functionality but at the same time filtering out semantically implausible sentences to facilitate the task of the human translators.
The results of this process are two small testing corpora, each consisting of 50 sentences in English, generated using the abstract syntax and the English concrete syntax of the DMI and the CSE grammar respectively.

The reference translations were obtained by asking two native speakers of Italian and Swedish to compare the original English sentences to their automatically translated counterparts and correct the latter.
In order to obtain more meaningful BLEU scores, participants were instructed to only make the minimal changes necessary to obtain, starting from the automatic tranlsations generated by our system, a set of grammatically and semantically correct translations.

\subsubsection{Experimental results}
